### 2015-09-11 paper.tex
Rearrange first couple introduction paragraphs

\section*{Introduction} Sandia National Laboratory (SNL) developed the
cross-flow turbine, Reference Model 2 (RM2), to provide an open-source point
design for marine hydrokinetic (MHK) energy research and development, including
its use as a test object for scaled model testing to generate power performance
data that can be used to validate open-source design tools. More details on the
reference modeling effort are provided in Neary et al. \cite{Neary2014}.

Dimensional analysis provides scaling laws that are used to upscale model test
data into performance and design information for a full-scale prototype turbine.
For hydrokinetic turbines, hydrodynamic similitude is achieved when the Reynolds
number, Froude number and tip-speed-ratio of the model and the full scale device
match. The Reynolds number is defined as
\begin{equation}
Re_L = \frac{UL}{\nu},
\label{eq:Re}
\end{equation}
where $U$ and $L$ are characteristic length scales, and $\nu$ is the fluid
kinematic viscosity. The Froude number is defined as
\begin{equation}
Fr = \frac{U}{\sqrt{gL}},
\label{eq:Fr}
\end{equation}
where $g$ is the gravitational acceleration, and the tip speed ratio is defined
as
\begin{equation}
\lambda=\frac{\omega R}{U_\infty},
\end{equation}
where $\omega$ and $R$ are the rotor angular velocity and radius.
For hydrokinetic turbine testing, $Re$, not $Fr$, is the dominant scaling
parameter. However, the Froude number (based on blade tip submergence) should
remain small (significantly below one). It is rare to achieve perfect similitude
for $Re$, but a threshold value should be exceeded, such that scale model
results can be extrapolated.

As a cross-flow turbine, the RM2 blades will experience large variations in
angles of attack as they rotate about their axis (``cross-flow turbine'' means
the axis of rotation is perpendicular to flow direction---either vertically or
horizontally). These variations become larger as the tip speed ratio decreases
\cite{Para2002}. Typically, the maximum angles of attack are sufficiently large
that the blades operate under dynamic stall, which is a complex unsteady process
and deviates significantly from static foil behavior, during part of the
rotation. Furthermore, the higher the solidity of a cross-flow turbine, the
lower the optimal tip-speed ratio at which it operates. Marine Hydrokinetic
(MHK) cross-flow turbines typically have higher solidity than cross-flow wind
turbines (VAWT), and hence operate at lower tip-speed ratios. Since MHK turbines
operate in a higher density fluid, unsteady dynamic effects related to the
blades' pitching motion and flow curvature also become more important when
compared to wind turbines.

>>>>

\section*{Introduction}

Sandia National Laboratories, in collaboration with the US Department of Energy
developed the Reference Model devices to provide modeling targets for advancing
marine hydrokinetic (MHK) energy technology \cite{Neary2014}. The Reference
Model 2 (RM2), a cross-flow turbine intended for river environments, was
initially designed by Barone et al. \cite{Barone2011} using the CACTUS vortex
line code \cite{Murray2011}. A scaled physical model of the RM2 was then
designed and built by the Saint Anthony Falls Laboratory (SAFL) to validate the
CACTUS predictions. Experimental performance measurements were significantly
lower than the numerical results \cite{Hill2014}, which was assumed to be caused
by the physical model's low Reynolds number
\begin{equation}
Re_L = \frac{UL}{\nu},
\label{eq:Re}
\end{equation}
where $U$ and $L$ are characteristic length scales, and $\nu$ is the fluid
kinematic viscosity.

As a cross-flow turbine, the RM2 blades experience large variations in angles of
attack as they rotate about their axis (``cross-flow turbine'' means the axis of
rotation is perpendicular to flow direction---either vertically or
horizontally). These variations become larger as the tip speed ratio
\begin{equation}
\lambda=\frac{\omega R}{U_\infty},
\end{equation}
where $\omega$ and $R$ are the rotor angular velocity and radius, decreases
\cite{Para2002}. Typically, the maximum angles of attack are sufficiently large
that the blades operate under dynamic stall, which is a complex unsteady process
and deviates significantly from static foil behavior, during part of the
rotation. Furthermore, the higher the solidity of a cross-flow turbine, the
lower the optimal tip-speed ratio at which it operates. Marine Hydrokinetic
(MHK) cross-flow turbines typically have higher solidity than cross-flow wind
turbines (VAWT), and hence operate at lower tip-speed ratios. Since MHK turbines
operate in a higher density fluid, unsteady dynamic effects related to the
blades' pitching motion and flow curvature also become more important when
compared to wind turbines.
